---
title: "Querying prescription drug claims"
author: "Benjamin Chan (chanb@ohsu.edu)"
date: "Monday, January 12, 2015"
output:
  ioslides_presentation:
    keep_md: yes
    smaller: yes
    widescreen: yes
---
  
  
## The problem
  
* You need to look for prescription drug claims for drug *XYZ*
* Prescription drug claims usually only have NDC codes to describe the prescription filled
    + Drug names are typically not on claims
    + Drug classes, if present, may cast too wide of a net
* There are ~150K NDC codes listed on [fda.gov](http://www.fda.gov/Drugs/InformationOnDrugs/ucm142438.htm)
* Going to [fda.gov](http://www.fda.gov/Drugs/InformationOnDrugs/ucm142438.htm) would be easy *except*...
    + [fda.gov](http://www.fda.gov/Drugs/InformationOnDrugs/ucm142438.htm) uses a different coding system than what appears on claims
* **How to query prescription drug claims for only certain drugs?**
  
  
## Step 0
  
If you can get a topic expert to give you a list of NDC codes to query for, then you're golden... Except

* What if there are new drugs in the same class on the market?
* What if there are new dosages for the same substance?
* What if there are new packagings for the same substance?
* What if there are new manufacturers for the same substance?
* What is the topic expert mistypes a code?

Would like to have a less error-prone process.


## Step 1

Import the NDC translation dataset into your analysis package.

```{r, echo=FALSE}
require(data.table, quietly=TRUE)
setInternet2()
```

It is maintained by me at [GitHub](https://github.com/benjamin-chan/NDCTranslation). Go ahead and fork it!
The Git repo has 2 branches

* `master` produces a [zipped data file](https://raw.githubusercontent.com/benjamin-chan/NDCTranslation/master/NDCTranslationTable.zip)

```{r, echo=FALSE, eval=FALSE}
url <- "https://raw.githubusercontent.com/benjamin-chan/NDCTranslation/master/NDCTranslationTable.zip"
f <- tempfile()
download.file(url, f)
con <- unz(f, "NDCTranslationTable.txt")
ndcMaster <- read.delim(con, stringsAsFactors=FALSE)
ndcMaster <- data.table(ndcMaster)
```

* `TextFile` produces a [tab-delimited data file](https://raw.githubusercontent.com/benjamin-chan/NDCTranslation/TextFile/NDCTranslationTable.txt)

```{r, eval=FALSE}
url <- "https://raw.githubusercontent.com/benjamin-chan/NDCTranslation/TextFile/NDCTranslationTable.txt"
ndcMaster <- fread(url, colClasses=rep("character", 22))
```

```{r, echo=FALSE}
readNDC <- function () {
  url <- "https://raw.githubusercontent.com/benjamin-chan/NDCTranslation/TextFile/NDCTranslationTable.txt"
  ndcMaster <- fread(url, colClasses=rep("character", 22))
  ndcMaster
}
t <- system.time(ndcMaster <- readNDC())
message(sprintf("This took %.2f minutes", t[3] / 60))
```


## Step 2

Go crazy with regular expressions to produce a drug lookup table.

Some of the more important columns

* `nonproprietaryname`
* `substancename`
* `routename`
* `pharm_classes`
* `ndcPackageFmt11` (11-digit NDC format found on claims; this will be you key variable)

```{r}
ndcLookupA <- ndcMaster[grep("angiotensin.+receptor.blocker", pharm_classes, ignore.case=TRUE)]
ndcLookupB <- ndcMaster[grepl("methadone|buprenorphine|naloxone", substancename, ignore.case=TRUE)]
ndcLookupC <- ndcMaster[grep("[a-z]tazone", nonproprietaryname, ignore.case=TRUE)]
```

## Step 3

Check.

```{r}
ndcLookupA[, .N, substr(pharm_classes, 1, 60)][order(N, decreasing=TRUE)]
ndcLookupA[grep("^increased", pharm_classes, ignore.case=TRUE),
           .N,
           substancename]
```

## Step 3

Check.

```{r}
ndcLookupB[, .N, substancename][order(N, decreasing=TRUE)]
ndcLookupB[, .N, routename][order(N, decreasing=TRUE)]
```

## Step 3

Check.

```{r}
ndcLookupC[, .N, nonproprietaryname][order(N, decreasing=TRUE)]
```


## Step 4

Do an inner join on the claims data to the drug lookup table.

```{r, eval=FALSE}
setkey(rxClaims, ndc)
ndcLookupA <- ndcLookupA[, ndc := ndcPackageFmt11]
setkey(ndcLookupA, ndc)
rxClaimsA <- merge(rxClaims, ndcLookupA)
```
