---
title: "Finding prescription drug claims"
author: "Benjamin Chan (chanb@ohsu.edu)"
date: "Monday, January 12, 2015"
output: ioslides_presentation
---


## The problem

* Prescription drug claims usually only have NDC codes
    + Drug names are usually not on claims
    + Drug classes may be cast too wide of a net
* There are ~150K NDC codes listed on the [fda.gov](http://www.fda.gov/Drugs/InformationOnDrugs/ucm142438.htm)
* Going to [fda.gov](http://www.fda.gov/Drugs/InformationOnDrugs/ucm142438.htm) would be easy *except*
    + [fda.gov](http://www.fda.gov/Drugs/InformationOnDrugs/ucm142438.htm) uses a different coding system than what appears on claims
* **How to query prescription drug claims for only certain drugs?**


## Step 0

If you can get a topic expert to give you a list of NDC codes to query for, then you're golden... Except

* What if there are new drugs on the market?
* What if there are new dosages for the same substance?
* What if there are new packagings for the same substance?
* What if there are new manufacturers for the same substance?


## Step 1

Grab the [NDC translation dataset](https://raw.githubusercontent.com/benjamin-chan/NDCTranslation/TextFile/NDCTranslationTable.txt).

It is maintained by me at [GitHub](https://github.com/benjamin-chan/NDCTranslation).


## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```

